---
title: 'Learn to R: Swirl Week 1'
author: "Jonathan Roberts"
date: "20 March 2017"
output:
  html_notebook: default
  pdf_document: default
  html_document: default
---

```
#{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

#Packages and Setup
Download and set-up the swirl courses ready to go

```{r}

#getwd()
#setwd("/Users/jonathanroberts/Documents/R/swirl")
#install.packages("swirl")
#packageVersion("swirl")
#library("swirl")
#ls()
#rm(list = ls())
#install_from_swirl("R Programming")
#swirl()
#dir()

```

#Swirl R Programming: section 1: Basic building blocks
To basic to bother writting the code.
```{r results = "hide"}
5+7

```

#Swirl R Programming: section 2: Workspace and files

```{r results = "hide"}
getwd()
ls()
x<-9
ls()
dir()

args(list.files)
old.dir<-getwd()
dir.create("testdir")
setwd("/Users/jonathanroberts/Documents/R/swirl/testdir")
file.create("mytest.R")
file.exists("mytest.R")
file.info("mytest.R")
file.rename("mytest.R", "mytest2.R")
file.copy("mytest2.R", "mytest3.R")
file.path("mytest3.R")
dir()
setwd(old.dir)

```

# Swirl R Programming: section 3: Basic Sequences

```{r results = "hide"}

1:20
pi:10
15:1
seq(1,20)
seq(0,10, by=0.5)
my_seq<-seq(5,10, length=30)
length(my_seq)

1:length(my_seq)
seq(along.with = my_seq)
seq_along(my_seq)

rep(0, times = 40)
rep(c(0,1,2), times=10)
rep(c(0,1,2), each=10)



```
#Swirl R Programming: section 4: Vectors 
Atomic Vectors contain 1 data type. Play around a bit with character and logical vectors.
```{r}
num_vect<-c(0.5, 55, -10, 6)
tf <- num_vect<1
num_vect >= 6

my_char <- c("My", "name","is")
my_char

paste(my_char, collapse = " ")
my_name <- c(my_char, "Jon")

my_name

paste(my_name, collapse = " ")
paste("Hello", "world!", sep = " ")
paste(1:3,c("X", "Y", "Z"), sep = "")
paste(LETTERS, 1:4, sep="-")

```

#Swirl R Programming: section 5: Missing Values

```{r}
x<-c (44, NA, 5, NA)
3*x
y<-rnorm(1000)
z<-rep(NA, 1000)
my_data<-sample(c(y,z), 100)

my_na<-is.na(my_data)

my_na
# testing what happens is you do ==na. This doesn't work as NA is not a value to put into the logical operator
#my_data == NA

#Count how many NA's in the data
sum(my_na)
my_data

0/0
Inf - Inf

```


#Swirl R Programming: section 6: Subsetting Vectors



```{r}

#so we are given a vector with 20 random numbers from normal dist and 20 NA's we'd create like this 
y<-rnorm(20)
z<-rep(NA, 20)
# not sure how you pick 20 from each but at random - the following won't do that i'm pretty sure! 
my_data<-sample(c(y,z), 40)

x
x[1:10]
x[is.na(x)]
y<-x[!is.na(x)]
y
y[y>0]
# you have to strip out the NA's first so that the logical operators will work hence
x[x>0]
# leaves all the Na's in
#we could do the operations together without using y though

x[!is.na(x) & x > 0]


#Now return a specific selection

x[c(3,5,7)]

x[0] # doesn't work as r indexes from 1 not 0...
x[3000] # it returns an NA - even though out of bounds!!!

x[c(-2,-10)] # which is same as
x[-c(2,10)]

vect <- c(foo=11, bar = 2, norf = NA)
vect
names(vect)
vect2 <- c(11, 2, NA)
names(vect2) <- c("foo","bar", "norf")
identical(vect,vect2)

vect["bar"]
vect[c("foo", "bar")]

```

#Swirl R Programming: section 7: Matrices and DataFrame

```{r}

my_vector <- 1:20
my_vector
dim(my_vector)
length(my_vector)
dim(my_vector) <- c(4,5)
dim(my_vector)
attributes(my_vector)
my_vector
class(my_vector)
my_matrix<-my_vector
?matrix

matrix(data = 1:20, nrow=4, ncol = 5)
matrix(1:20, nrow=4, ncol = 5)
my_matrix2<-matrix(1:20, 4, 5)

identical(my_matrix, my_matrix2)
patients <- c("Bill", "Gina", "Kelly", "Sean")
cbind(patients, my_matrix) # note this screws up the Matrix as you've added character vector into a numeric matrix!!

my_data <- data.frame(patients, my_matrix)
my_data
class(my_data)
cnames<-c("patient", "age", "weight", "bp", "rating", "test")
colnames(my_data)<-cnames
my_data

```


